An Open System for\\ Managing Short Programming Exercises

### 2013-08-09 paper/talk.tex
grading questions?
tightening up to fit in 7 minutes

%%%%%%%%%%%%%%%%%%%%%%%% Outline %%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Outline}

\begin{itemize}
  \item Background
  \item Description of CloudCoder
  \item Preliminary Experiences
  \item Questions
\end{itemize}

@@

\begin{itemize}
  \item Many students struggle in introductory programming courses
  \item Mastery of basic concepts and techniques is essential for success
  \item Short programming exercises ({\it {\` a} la} CodingBat):
  \begin{itemize}
    \item Students write or complete a short (5-15 line) program
    \item Interface is typically web-based
    \item Correctness is judged with a series of test cases
    \item Student receives immediate feedback
  \end{itemize}
  \item Exercises are useful for:
  \begin{itemize}
    \item Skills development
    \item Self-tests to accompany reading
    \item Extra practice
  \end{itemize}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%% Short Exercises %%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Short Exercises}

\begin{itemize}
  \item Analogy: learning programming vs.\ learning to play a musical instrument
  \item Larger assignment (like SIGCSE {\em Nifty Assignments}) :: Recital piece
  \item Short exercise :: Playing a scale, an {\' e}tude
\end{itemize}

@@

\begin{itemize}
  \item There are lots of programming exercise systems already:
  \begin{itemize}
    \item Free\footnote{As in beer}: CodingBat, Codecademy
    \item Commercial: CodeLab (TuringsCraft), MyProgrammingLab (Pearson)
    \item {\em Do we really need another one?}
  \end{itemize}
  \item We did not know of one that
  \begin{itemize}
    \item {\em Encouraged users to create and share exercises}
    \item Supported a range of languages (including C/C++)
    \item Was open source
    \item Collected rich data about how students work
  \end{itemize}
  \item We created a new system, called CloudCoder, to fill this niche
    \begin{itemize}
    \item \url{http://cloudcoder.org}
    \end{itemize}
\end{itemize}

@@

\begin{itemize}
  \item The impact of a pedagogical technique or curriculum resource
        depends on how many students can access it
  \item Systematically eliminating barriers to adoption and distribution maximizes impact
  \begin{itemize}
    \item Requiring payment is a barrier
    \item Restrictive licenses are a barrier
    \item Control by a central authority is a barrier
  \end{itemize}
  \item Ideal situation: resources are available to {\bf anyone} at {\bf any time} for {\bf any purpose}
%  \item Case in point...
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%% Our philosophy %%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Our philosophy}

\begin{itemize}
  \item Make programming exercises written by many people available
        for whatever uses are appropriate
  \begin{itemize}
    \item Our need: exercises to assign in the context of a course
    \item Related work: PCRS (Zingaro et.\ al., SIGCSE 2013), in-class
          exercises to support peer instruction
      \begin{itemize}
      \item Summer 2013: PCRS can import exercises from the CloudCoder exercise repository!
      \end{itemize}
    \item We can imagine lots of other uses: for example, Codecademy-style
          self-learning
  \end{itemize}
  \item Make the {\em sharing} of exercises a subject of research
  \begin{itemize}
    \item What kinds of exercises are most easily reused?
    \item How can exercises be classified and organized?
  \end{itemize}
\end{itemize}

@@

\begin{itemize}
  \item Open source (AGPL v3)
  \item Web-based user interface
    \begin{itemize}
    \item 100\% HTML/CSS/JavaScript, no plugin required
    \end{itemize}
  \item Supports exercises in Java, Python, Ruby, and C/C++
  \item Collects fine-grained edit sequence of each student's work
  \item Repository of freely-redistributable (CC-BY-SA) exercises:
        \url{https://cloudcoder.org/repo}
\end{itemize}

@@

@@

\begin{itemize}
  \item Requires two Linux servers to host (one network-facing)
    \begin{itemize}
    \item Relatively easy to install
    \item Known to work: 60+ concurrent users with modest hardware
    \item Load testing: with 700+ concurrent users
          on moderate server hardware
      \begin{itemize}
      \item Compiling and testing submissions can require significant
            CPU resources, but is easy to scale up by adding more
            build servers
      \end{itemize}
    \item Cost to host per course-semester: perhaps \$1-\$3 per student
          (e.g., using Amazon EC2)
    \end{itemize}
  \item {\bf We can host CloudCoder for your course!}
    \begin{itemize}
    \item We have a grant from Amazon to cover hosting costs for
          a small number of institutions/courses
    \item {\bf Talk to us if you are interested!}
    \end{itemize}
\end{itemize}

@@

\begin{itemize}
\item Are short exercises useful?
  \begin{itemize}
  \item Do you use short exercises in your courses currently?
    \begin{itemize}
    \item If so, how?
    \item If not, would you be interested in starting to use them?
    \end{itemize}
  \item What experiments should we do to measure effect on outcomes?
  \end{itemize}
\item How should we most effectively incorporate short exercises into a course?
  \begin{itemize}
  \item How do we know what exercises should be assigned?
  \item In what order should the exercises be assigned?
  \item What mix of short exercises vs.\ larger assignments?
  \item How to motivate students to attempt exercises?
  \end{itemize}
\end{itemize}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%% Outline %%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{frame}{Outline}
%
%\begin{itemize}
%  \item Background
%  \item Description of CloudCoder
%  \item Preliminary Experiences
%  \item Questions
%\end{itemize}
%
%\end{frame}

@@

\begin{itemize}
  \item Many students struggle in introductory programming courses
  \item Mastery of basic concepts and techniques is essential for success
  \item Short programming exercises ({\` a} la CodingBat):
  \begin{itemize}
    \item Web-based UI
    \item Students write or complete a short (5-15 line) function or program
    \item Correctness is judged with a series of test cases
    \item Student receives immediate feedback
  \end{itemize}
  \item Useful for:
  \begin{itemize}
    \item Skills development
    \item Self-tests to accompany reading
    \item Extra practice
  \end{itemize}
\end{itemize}

@@

\begin{itemize}
  \item There are lots of programming exercise systems already:
  \begin{itemize}
    \item Free: CodingBat, PracticeIt!, Codecademy
    \item Commercial: CodeLab (TuringsCraft), MyProgrammingLab (Pearson)
    \item {\em Do we really need another one?}
  \end{itemize}
  \item We did not know of one that
  \begin{itemize}
    \item Encouraged users to create and share {\bf freely-redistributable} exercises
    \item Supported a range of languages (including C/C++)
    \item Was open source
    \item Collected rich data about how students work
  \end{itemize}
  \item We created a new system, called CloudCoder, to fill this niche
    \begin{itemize}
    \item \url{http://cloudcoder.org}
    \end{itemize}
\end{itemize}

@@

@@

\begin{itemize}
  \item Open source (AGPL v3)
  \item Supports exercises in Java, Python, Ruby, and C/C++
  \item Web-based user interface
    \begin{itemize}
    \item 100\% HTML/CSS/JavaScript, no plugin required
    \end{itemize}
  \item Repository of freely-redistributable (CC-BY-SA) exercises:
        \url{https://cloudcoder.org/repo}
    \begin{itemize}
    \item We encourage users to contribute, but do not require them to
    \end{itemize}
  \item Collects fine-grained edit sequence of each student's work
\end{itemize}

@@

\begin{itemize}
  %\item The impact of a pedagogical technique or curriculum resource
  %      depends on how many students can access it
  \item Goal: make programming exercises available to {\bf anyone} at {\bf any time} for {\bf any purpose}
  \begin{itemize}
    \item No payment
    \item No restrictive licenses
    \item No central control
  \end{itemize}
  \item Exercises can be used in many contexts:
  \begin{itemize}
    \item Our need: assign exercises in the context of a course
    \item In-class exercises to support peer instruction
          (PCRS, Zingaro et.\ al., SIGCSE 2013)
      \begin{itemize}
      \item Summer 2013: PCRS can import exercises from the CloudCoder exercise repository!
      \end{itemize}
    \item Codecademy-style self-learning?
  \end{itemize}
\end{itemize}

@@

\begin{itemize}
  \item Requires two Linux servers to host (one network-facing)
    \begin{itemize}
    \item Relatively easy to install
    \item Known to work: 60+ concurrent users with modest hardware
    \item Load testing: 700+ concurrent users
          on moderate server hardware
      \begin{itemize}
      \item Compiling and testing submissions can require significant
            CPU resources, but are easy to scale up by adding more
            build servers
      \end{itemize}
    \item Cost to host per course-semester: perhaps \$1-\$3 per student
          (e.g., using Amazon EC2)
    \end{itemize}
  \item {\bf We can host CloudCoder for your course!}
    \begin{itemize}
    \item We have a grant from Amazon to cover hosting costs for
          a small number of institutions/courses
    \item {\bf Talk to us if you are interested!}
    \end{itemize}
\end{itemize}

@@

\begin{itemize}
\item Are short exercises useful?
  \begin{itemize}
  \item Do you use short exercises in your courses currently?
    \begin{itemize}
    \item If so, how?
    \item If not, would you be interested in starting to use them?
    \end{itemize}
  \item What experiments should we do to measure effect on outcomes?
  \end{itemize}
\item How to incorporate short exercises into a course most effectively?
  \begin{itemize}
  \item How do we know what exercises should be assigned?
  \item In what order should the exercises be assigned?
  \item What mix of short exercises vs.\ larger assignments?
  \item How to motivate students to attempt exercises?
    \begin{itemize}
    \item Should they be graded?
    \end{itemize}
  \end{itemize}
\end{itemize}
